---
title: "CHL8010: Statistical Programming and Computation in Health Data"
author: "Aoqi Xie"
date: "2024-09-30"
format: pdf
---

```{r}
#| label: load-packages
#| include: false
# source("../.Rprofile")
library(here)
library(dplyr)
library(ggplot2)
library(tidyr)
library(usethis)
library(purrr)
library(countrycode)
library(corrplot)
```


## Perfect your GitHub repo

Some of you may still need to organize your GitHub repo. Use this time to do that. When you are confident with your repo, let me know -- I will try to reproduce your code. 

Your final data should have the following variables (you might have slightly different variable names).
```{r}
finaldata <- read.csv(here("data", "primaryanalysis_data.csv"), 
                      header = TRUE)
names(finaldata)
```

Observations from Canada should look like this...
```{r}
finaldata %>%
  dplyr::filter(country_name == "Canada")
```

Observations from Ecuador should look like this...
```{r}
finaldata %>%
  dplyr::filter(country_name == "Ecuador")
```

## Exploratory data analysis
Use the rest of the class time to explore the final data that will be used for analysis starting next week. At the end of the class, write a summary of your findings and push your **Quarto document (pdf)** to your repo. 

1. Understand the structure of the data
From the results below, there are a total of 5320 observations with 21 variables (8 integer types, 10 numerical types, and 3 character types). We can also see some missingness in the data set, which should be carefully taken care of.
```{r}
str(finaldata)                       
summary(finaldata) 
```

2. Understand the distribution and skewness of the outcome data (total death from the conflict data)
```{r}
ggplot(finaldata, aes(x = Year, y = totdeath)) +
  geom_histogram(stat = "identity", position = "dodge") +
  facet_wrap(~ region, nrow = 9, ncol = 2) +
  labs(title = "Histogram of total deaths by Year and Region",
       x = "Year",
       y = "Totdeath") +
 theme_minimal(base_size = 10) +  # Increase base font size for better visibility
  theme(
    strip.text = element_text(size = 10),  # Size of facet labels
    plot.title = element_text(size = 12, face = "bold"),  # Title size
    axis.title = element_text(size = 12),  # Axis title size
    axis.text = element_text(size = 10)  # Axis text size
  )
```
3. Understand the correlation among the numeric variables. We can see that the mortality rates are highly correlated, and moderate correlation observed among ``gdp1000``, ``OECD'', and ``popdens''.
```{r}
correlation_matrix <- cor(finaldata %>% select_if(is.numeric), use = "complete.obs")  # Select only numeric columns
corrplot(correlation_matrix, method = "circle")  # Visualize correlation matrix
```
